\documentclass[manual.tex]{subfiles} 
\begin{document}
\section{Introduction}\label{sec:intro}

This document is the manual for SMAC~\cite{HutHooLey11-SMAC} (an acronym for \emph{Sequential Model-based Algorithm Configuration}). SMAC aims to solve the following \emph{algorithm configuration} problem: Given a binary of a parameterized algorithm $\mathcal{A}$, a set of instances $\mathcal{S}$ of the problem $\mathcal{A}$ solves, and a performance metric $m$, find parameter settings of $\mathcal{A}$ optimizing $m$ across $\mathcal{S}$.

In slightly more detail, users of SMAC must provide:
\begin{itemize}
\item a parametric algorithm $\mathcal{A}$ (an executable to be called from
the command line), 
\item a description of $\mathcal{A}$'s parameters $\theta_1,\dots,\theta_n$ and their domains $\Theta_1, \dots, \Theta_n$, 
\item a set of benchmark instances, $\Pi$, and
\item the objective function with which to measure and aggregate algorithm preformance results.
\end{itemize}

SMAC then executes algorithm $\mathcal{A}$ with different \emph{parameter configurations} (combinations of parameters 
$\langle{}\theta_1,\dots,\theta_n\rangle{} \in \Theta_1 \times \cdots \times \Theta_n$, on instances $\pi \in \Pi$),
searching for the configuration that yields overall best performance across the benchmark instances under the supplied objective. For more details please see~\cite{HutHooLey11-SMAC}; if you use SMAC in your research, please cite that article. It would also be nice if you sent us an email -- we are always interested in additional application domains.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{License}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This version of SMAC is released under the AGPLv3, please read \texttt{LICENSE-AGPLv3.txt} for more information. For other license options please e-mail Frank Hutter. Note that versions of SMAC prior to v2.10 were released under a different version.

\subsection{System Requirements}

SMAC itself requires only Java 7 or newer to run. SMAC is primarily intended to run on Unix like platforms, but now includes start up scripts so that it can run on Windows. You should be able to follow the commands exactly as laid out in this document but note that you may need to change the ending of the script to \texttt{.bat} (for instance \texttt{smac} becomes \texttt{smac.bat}), additionally you may want to substitute backslashes instead of forward slashes for paths.

Most of the included scenarios (in \texttt{./example\_scenarios/} require ruby and Linux 32-bit libraries to run. The scenarios in \texttt{./example\_scenarios/branin/} require only python and are cross platform. The saps scenario should work on Windows, Linux and Mac OS X (no BSD or Solaris).

\subsection{Version}
This version of the manual is for SMAC \input{versionfull}$\!\!$.
\\
\subfile{githashes.tex}



%\subsection{Included Example Scenarios}
%\note{FH: TODO after we decide what to include}
\end{document}